---
title: "Reproducible Research PA_1"
author: "Ian Yang"
date: "January 4, 2018"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading and preprocessing the data
```{r}
data = read.csv("activity.csv", header = T)
head(data)
```

#What is mean total number of steps taken per day? (ignore missing value)
```{r}
data_omit = na.omit(data)
head(data_omit)
total_day = aggregate(x = data_omit[c("steps")], FUN = sum, by = list(Group.date = data_omit$date))
```
###Histogram of total steps per day
```{r}
hist(total_day$steps, breaks=10, xlab = "Steps", main = "Total Steps per Day", xlim = c(0,25000))
```
###Mean/Median of steps taken per day
```{r}
mean_day = mean(total_day$steps)
mean_day
median_day = median(total_day$steps)
median_day
```
###What is the average daily activity pattern

```{r}
interval = aggregate(x = data_omit[c("steps")], FUN = mean, by = list(interval = data_omit$interval))
library(ggplot2)

ggplot(interval, aes(x = interval, y = steps)) + geom_line(color = "green", size = 1) +
  labs(title = "Average Steps Taken per Day", x = "Interval", y = "Avg. Steps per day")
```
###Interval with the max. number of steps
```{r}
interval[which.max(interval$steps), ]
```


#Imputing missing values
###Count missing values
```{r}
sum(is.na(data))
```
###Filling in all of the missing values by using mean
```{r}
data2 = data
nas = is.na(data2$steps)
avg_interval = tapply(data2$steps, data2$interval, FUN = mean, na.rm = TRUE, simplify = TRUE)
data2$steps[nas] = avg_interval[as.character(data2$interval[nas])]

total_day2 = aggregate(x = data2[c("steps")], FUN = sum, by = list(Group.date = data2$date))

hist(total_day2$steps, breaks=10, xlab = "Steps", main = "Total Steps per Day", xlim = c(0,25000))
mean_day = mean(total_day2$steps)
mean_day
median_day = median(total_day2$steps)
median_day
```
###Conclusion: After filling the missing value, I found that both mean and median of total steps per day are the same

#Are there differences in activity patterns between weekdays and weekends?
###Create New factor variables weektype
```{r}
library(magrittr)
library(dplyr)
head(data2)
data2$date = as.Date(data2$date, format = "%Y-%m-%d")
data2 = data2%>%
  mutate(weektype = ifelse(weekdays(data2$date) =="Saturday"|weekdays(data2$date) =="Sunday", "Weekend", "Weekday"))

head(data2)
data_weekday = data2[which(data2$weektype == "Weekday"), ]
head(data_weekday)
data_weekend = data2[which(data2$weektype == "Weekend"), ]
head(data_weekend)

interval_weekday = aggregate(x = data_weekday$steps, FUN = mean, by = list(interval = data_weekday$interval))
interval_weekday = interval_weekday%>%
  mutate(weektype = "Weekday")
interval_weekend = aggregate(x = data_weekend$steps, FUN = mean, by = list(interval = data_weekend$interval))
interval_weekend = interval_weekend%>%
  mutate(weektype = "Weekend")

colnames(interval_weekday) = c("interval", "avg_steps", "weektype")
colnames(interval_weekend) = c("interval", "avg_steps", "weektype")
interval2 = rbind(interval_weekday, interval_weekend)
head(interval2)
```
###Plot
```{r}
plot = ggplot(interval2, aes(x = interval, y = avg_steps, color = weektype)) + 
  geom_line() + labs(title = "Average Steps Taken per Day", x = "Interval", y = "Avg. Steps per day") +
  facet_wrap(~weektype, ncol = 1, nrow = 2)

print(plot)
```

###The test object is more active in the morning during weekday but overall walk more during the weekend. It's intuitively makes sense because people usually go to work during weekday and have more spare time during weekend.

